---
title: "Species with celebrity names"
author: ""
output:
  # bookdown::word_document2:
  bookdown::pdf_document2:
  toc: false
number_sections: false
fig_caption: true
highlight: "monochrome"
---

```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
library(knitr)
library(here)
```

# Supplementary methods

We fit Bayesian hierarchical generalized linear mix effects models (GLMMs) to assess the effect of species being named after celebrities on Wikipedia view counts.
These GLMMs were constructed with random intercepts to account for the paired nature of species matched based on TODO.

We considered two forms of observation error:
(1) Student-t observation error on the (natural) log of average species daily views on Wikipedia, and
(2) negative binomial ('NB2' [REF]) observation error on total number of species views.

The first option models a continuous value (log of average species daily views) with a heavy-tailed observation error distribution.
This option is also closest to the preregistered methodological approach.
We were unable to use a simple Gaussian error distribution as noted in the preregistration because of the heavy-tailed nature of the observations (the confidence intervals would have been overly precise and the parameters biased by rare outlying values).
The Student-t model necessitated excluding (TODO) species pairs where one or the other had zero views.

The second option is a count distribution where the observation variance grows quadratically with the expected value.
This maintains the integer nature of the data generating process and allows for zeros but may not sufficiently account for outliers.

Starting with the Student-t model, our model was of the form:
\begin{align}
\log(y_i) &\sim \mathrm{Student-t} \left( \nu, \mu_i, \sigma \right),\\
\mu_i &= \alpha + \alpha_j + \delta_s + \beta C_i + \beta_j C_i,
\end{align}
where $y_i$ represents the response variable (average species daily views) for observation $i$,
$\mu$ represents the expected value (mean), and $\sigma$ represents a scale parameter.
We assumed the observation error was drawn from a heavy-tailed Student-t distribution with $\nu = 4$ to account for occasional outliers.
We chose $\nu = 4$ based on posterior predictive checks.
The parameter $\alpha$ is a global intercept,
$\alpha_j$ is a taxa-specific (indexed by $j$) intercept, and
$\delta_s$ is a matched species ID (indexed by $s$) intercept.
The coefficient $\beta$ is a global effect of celebrity status ($C$, $C = 0$ if a non-celebrity, $C = 1$ if a celebrity) and
$\beta_j$ is a taxa-specific effect of celebrity status.

We constrained the taxa-specific intercepts $\alpha_j$, taxa-specific slopes $\beta_j$, and the celebrity/species-specific intercepts $\delta_s$ according to normal distributions:
\begin{align}
\alpha_{j} &\sim \mathrm{Normal} \left(0, \tau_{\alpha}^2 \right),\\
\beta_{j} &\sim \mathrm{Normal} \left(0, \tau_{\beta}^2 \right),\\
\delta_{s} &\sim \mathrm{Normal} \left(0, \tau_{\delta}^2 \right).
\end{align}

We placed the following weakly informative priors on the estimated parameters:
\begin{align}
\beta &\sim \mathrm{Normal \left(0, 1 \right)},\\
\alpha &\sim \mathrm{Normal \left(0, 5 \right)},\\
\sigma &\sim \mathrm{Student-t \left(3, 0, 2.5 \right)},\\
\tau_{\alpha} &\sim \mathrm{Student-t \left(3, 0, 2.5 \right)},\\
\tau_{\beta} &\sim \mathrm{Student-t \left(3, 0, 2.5 \right)},\\
\tau_{\delta} &\sim \mathrm{Student-t \left(3, 0, 2.5 \right)}.
\end{align}

We placed an $\mathrm{LKJ}(1)$ prior on the correlation between $\beta_{j}$ and $\alpha_{j}$ as is the default in the package brms (REF).

The negative binomial model was the same except for the main equations:
\begin{align}
y_i &\sim \mathrm{NB2} \left( \mu_i, \phi \right),\\
\log (\mu_i) &\sim \alpha + \alpha_j + \delta_s + \beta C_i + \beta_j C_i,
\end{align}
where $y_i$ represents counts of page views.
We placed a $\mathrm{Student-t} \left(3, 0, 5 \right)$ prior on the shape parameter $\phi$.

We constructed the Stan (REF) models with the package brms (REF).
For each fitted model, we used brms and cmdstanr (REF) to sample 2000 iterations across 4 chains with 1000 iterations of warm-up per chain with the Stan No U-Turn Sampler.
We assessed convergence visually with traceplots and assuring that ESS (effect of sample size) was $>$ 100 and Rhat (the scale reduction factor) was < 1.01 for all parameters.
We assessed the ability for our probabilistic models to generate our observed data via density plots of posterior predictive checks.

To assess the degree to which the taxa with the most data were affecting inference for the data with fewer data (due to the hierarchical structure) and as a check that our Bayesian model with priors was not unduly influencing our inference,
we fit alternative versions of the Student-t model where each taxon was fit independently with (restricted) maximum likelihood and without priors.
In this case, each model took the form:
\begin{align}
\log(y_i) &\sim \mathrm{Student-t} \left( \nu, \mu_i, \sigma \right),\\
\mu_i &= \alpha + \delta_s + \beta C_i,\\
\delta_{s} &\sim \mathrm{Normal} \left(0, \tau_{\delta}^2 \right),
\end{align}
where $\alpha$ and $\beta$ are unique to each taxa.
This version retained the random intercept for $\delta_i$ to account for the paired nature of the data.

# Supplementary results

Overall, the NB2 and Student-t models provided similar inference.
However, based on slightly better posterior predictive checks, generally smaller effect sizes, and consistency with the preregistered study design we moved forward with the Student-t model.

# Supplementary figures

```{r dot-line-st4, fig.cap="Multiplicative effect on average daily Wikipedia views of a species being named after a celebrity compared to a similar species not named after a celebrity. This version uses a heavy-tailed Student-t observation error. Dots represent medians and thick and thin lines represent 50\\% and 95\\% credible intervals of the posterior. Colour lines represent definitions of a celebrity at increasing thresholds for average daily view counts. For example, the 100 threshold only includes species pairs where the celebrity Wikipedia page had at least 100 average daily views.", out.width="4in"}
f <- here("figs/dot-line-1-1000-st4.png")
include_graphics(f)
```

```{r dot-line-nb2, fig.cap="Same as previous figure but with negative binomial observation error.", out.width="4in"}
f <- here("figs/dot-line-1-1000-nb2.png")
include_graphics(f)
```

```{r dot-line-st4-ind, fig.cap="Same as previous figures but from independent models fit for each taxon using (restricted) maximum likelihood and without priors. Dots represent median effects and thick and thin lines represent 50\\% and 95\\% confidence intervals.", out.width="4in"}
f <- here("figs/independent-sdmTMB-models-student-4.png")
include_graphics(f)
```

```{r ppc-st4, fig.cap="Posterior predictive check of the Student-t model. Black line represents the density of the observed data. Blue lines represent the density of 25 random draws from the posterior predictive distribution. The panels represent the four threshold for defining a celebrity. We have cut off the x-axes to focus on the center of the distribution. At lower thresholds (1 and 10) we see that the observed data have heavier tails than the predictive data. This becomes much less apparent at thresholds of 100 or 1000.", out.width="6in"}
f <- here("figs/st4-ppcheck.png")
include_graphics(f)
```

```{r ppc-nb2, fig.cap="Same as previous figure but for the NB2 observation model.", out.width="6in"}
f <- here("figs/nb2-ppcheck.png")
include_graphics(f)
```

```{r violin-zoom, fig.cap="Visualization of raw data. The y-axis shows the ratio of average daily views for celebrity vs. non-celebrity species. The x-axis shows the average celebrity species page views. The red lines are robust linear regression for visualization purposes. We do not observe any strong patterns.", out.width="6in", eval=FALSE}
f <- here("figs/raw-dat-rlm.png")
include_graphics(f)
```

```{r raw-rlm, fig.cap="A visualization of the raw data distribution. Shown is the distribution of the ratio of average celebrity vs. non-celebrity species views across the four thresholds of defining a celebrity. The horizontal line is the median and the red outline is a violin (density) plot. The y-axes are truncated to focus on the center of the distribution.", out.width="6in"}
f <- here("figs/raw-dat-violin.png")
include_graphics(f)
```

\clearpage

```{r nb2-prob-table, eval=FALSE}
linesep <-  c('', '', '', '', '', '\\addlinespace')
tb <- readRDS(here("data-generated/nb2-probs.rds"))
knitr::kable(tb, digits = 2L, booktabs = TRUE, linesep = linesep,
  caption = "NB2 posterior"
  )
```

```{r st4-prob-table}
linesep <-  c('', '', '', '', '', '\\addlinespace')
tb <- readRDS(here("data-generated/nb2-probs.rds"))
knitr::kable(tb, digits = 2L, booktabs = TRUE, linesep = linesep,
  caption = "Posterior summary statistics for the hierarchical Student-t model. `CI' = credible interval. `Prob. effect > 1' is the posterior probability density above one. I.e., the probability of a positive effect of a species being named after a celebrity on a species' average daily Wikipedia views.", col.names = c("Celeb. threshold", "Taxon", "Prob. effect > 1", "Lower 95% CI", "Median", "Upper 95% CI"))
```
